Phase Noise Analysis in an Integer-N PLL

### 2025-05-08 Examples/Example3_4_SimulationandEstimationofLoopSettlingTimes/article/article.qmd
Finishing provisory reports.

---
title: Estimation of Fractional-N PLL Linearity Limits and Cycle Slipping
subtitle: Case Study - Example 3.3 from the Reference Material
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    This report explores the linearity limits of Phase-Frequency Detectors (PFDs) in Phase-Locked Loops (PLLs) and the onset of cycle slipping, a nonlinear phenomenon that occurs when the phase error exceeds the detector's effective operating range. The analysis focuses on tristate PFDs commonly used with charge pumps, emphasizing their functional behavior and limitations. A practical methodology for estimating the linear range through simulation is presented, along with a discussion on how exceeding this range can lead to loss of phase tracking and degraded loop performance. The consequences of cycle slipping, including increased jitter and potential loss of lock, are also examined.
keywords: 
    - Fractional-N PLL
    - Linearity Limit
    - Cycle Slipping
    - Nonlinear Dynamics
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

>>>>

---
title: Estimation of Fractional-N PLL Linearity Limits and Cycle Slipping
subtitle: Case Study - Example 3.4 from the Reference Material
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    This report explores the linearity limits of Phase-Frequency Detectors (PFDs) in Phase-Locked Loops (PLLs) and the onset of cycle slipping, a nonlinear phenomenon that occurs when the phase error exceeds the detector's effective operating range. The analysis focuses on tristate PFDs commonly used with charge pumps, emphasizing their functional behavior and limitations. A practical methodology for estimating the linear range through simulation is presented, along with a discussion on how exceeding this range can lead to loss of phase tracking and degraded loop performance. The consequences of cycle slipping, including increased jitter and potential loss of lock, are also examined.
keywords: 
    - Fractional-N PLL
    - Linearity Limit
    - Cycle Slipping
    - Nonlinear Dynamics
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

### 2025-05-08 Examples/Example3_5_SystemPhaseNoiseCalculation/article/article.qmd
Finishing provisory reports.

---
title: Integer-N PFD/CP-based PLL - System-Level simulation
subtitle: Laboratory 01 - Week 01-09
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    Este trabalho apresenta duas simulações para um Integer-N PFD/CP-based PLL a nível de sistema. Uma simulação é no domínio do tempo e a outra no domínio da fase. Foram estabelecidas restrições para a operação e são apresentados os passos até a concepção dos sistemas para a simulação. Por fim, os resultados são analisados quanto ao tempo de resposta e comparados ao que foi previsto com base na teoria.
keywords: 
  - Phase-locked loop
  - Sistema de controle
  - PLL Integer-N
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

@@

# Introduction
A `{python} format(f0)`Hz–`{python} format(f)`Hz synthesizer with a step size of `{python} format(step)`Hz is required, as the PLL is an Integer-N, $f_\text{ref}=$ `{python} format(fref)`Hz, a charge pump with `{python} format(I_CP)`A output current, and a VCO (operating from a `{python} format(V_sup)`V supply) are available. Designing a integer-N synthesizer with a loop bandwidth of `{python} format(f_3dB)`Hz using these components.

## VCO

For the VCO, if the system is fed with a `{python} format(V_sup)`V supply and the tuning range is `{python} format(tunning_range)`Hz, so the VCO constant can be $K_\text{VCO}=$ `{python} format(Kvco)`Hz/V. And initially the nominal frequency is arbitrary selected as $f_\text{VCO}=$ `{python} format(fvco)`Hz. With that nominal frequency, the division ratio will be $N=$ `{python} N`.

@@

In addition, since the charge pump current is known, the charge pump constant can be determined as $K_\theta=$ `{python} format(Ktheta)`A/rad.

@@

As the ration of reference frequency and bandwidth is $f_\text{ref}/f_{3\text{dB}}=$ `{python} round(fref_f3dB, 0)`, i.e. $f_\text{ref}\gg f_{3\text{dB}}$, the continuous-time analysis can be used.

This section is responsible to present the design of the loop filter. First of all, it's needed to determine the value of natural frequency,

@@

Choosing a damping constant, $\zeta=$ `{python} zeta`, $f_n=$ `{python} format(f_n)`Hz.

The first capacitor, $C_1$, can be calculated as

@@

So this capacitor has a value of, $C_1=$ `{python} format(C_1)`F.

The first resistor (and the single one) used in the second-order loop filter will can be obtained as follows,

@@

The value of that first resistor is $R_1=$ `{python} format(R_1)`$\Omega$. And the second capacitor, $C_2$, is obtained as one-tenth of $C_1$, getting the value of $C_2=$ `{python} format(C_2)`F.

# Phase noise

The intention of this report is to present analysis of phase noise in a integer-N PLL already designed.

@@

First of all, the VCO phase noise is supposed as being The VCO has a phase noise of $PN_\text{VCO}=$ `{python} pn_vco` dBc/Hz at a $\Delta f=$ `{python} format(offset_vco)`Hz offset (it bottoms out at $PN_{\text{VCO}_0}=$ `{python} noise_floor` dBc/Hz), and the charge pump puts out a noise current of $i_{n_\text{CP}}=$ `{python} format(I_noise_CP)`A/$\sqrt{\text{Hz}}$. Ignoring  PFD, divider, and reference noise sources.

Assuming that the VCO follows the 20 dB/dec rule it's possible to come up with a linear expression for the phase noise of the VCO based on

@@

So it's needed to get the parameters $C$ and $D$. Since the above equation must be linear at the logaritmic view, the parameter $C$ can be obtained as follows

@@

Thus $C=$ `{python} round(C_param, 1)` $\text{rad}^4/\text{Hz}^2$. The parameter $D$, for its time, can be calculated as

@@

And it results in $D=$ `{python} format_e(D_param)` $\text{rad}^2/\text{Hz}$. Thus the equation previously presented becomes

@@

The output noise current from the charge pump can be input-referred by dividing by $K_\theta$

@@

Using the above equation it results in $\varphi_\text{CP}=$ `{python} format(phi_CP)`$\text{rad}/\sqrt{\text{Hz}}$.

To calculate the noise provided by the second-order loop filter, it's needed to analyse that

The noise voltage develops a current flowing through the series combination of $C_1$ , $C_2$ , and $R$,

@@

Similarly, the analysis can be done for the voltage, i.e.,

@@

As the current noise is provided by the resistor, $i_n=\sqrt{4kT/R}$. To analyze the system noise properly, it's needed to refer the noise to the input,

@@

Now, it's possible to get the transfer function of the noise due to the PFD, charge pump, divider and loop filter (everything referred to the input) as follows

$$
\begin{split}
\frac{\varphi_\text{no}(s)}{\varphi_\text{nI}(s)}&=\frac{\frac{F(s)K_\text{VCO}K_\theta}{Ns}}{1+\frac{F(s)K_\text{VCO}K_\theta}{Ns}}\cdot N\cdot\frac{s}{s}\implies\\
&=\frac{F(s)K_\text{VCO}K_\theta}{s+\frac{F(s)K_\text{VCO}K_\theta}{N}}
\end{split}
$$
as the PSD output noise, $\varphi_\text{no}$, is obtained before the divider, the function could be multiplied by $N$ (i.e. divided by $N^{-1}$).

For a PLL with a second-order loop filter,

@@

Substituting the values, it results in

@@

display(Markdown(f"$$\\varphi_\\text{{no}}(s) = {latex_str}\\cdot {format_raw(phi_CP, 0)}$$ {{#eq-filter-transfer-function}}"))

Therefore, to plot phase noise in dBc/Hz, we take

@@

>>>>

---
title: Phase Noise Analysis in an Integer-N PLL
subtitle: Case Study - Example 3.5 from the Reference Material
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    This report presents the design and phase noise analysis of an Integer-N Phase-Locked Loop (PLL) frequency synthesizer. The system is specified to operate within a defined frequency range and step size, utilizing a voltage-controlled oscillator (VCO), charge pump, and loop filter. The loop filter is designed using continuous-time analysis, justified by the high ratio between the reference frequency and loop bandwidth. Each component is characterized to enable accurate modeling of the PLL’s behavior. Phase noise contributions from the VCO, charge pump, and passive components in the loop filter are identified and analyzed. The VCO noise is modeled based on its offset and floor characteristics, while noise from the charge pump and loop filter is referred to the input using appropriate transfer functions. The resulting model enables the calculation and visualization of the phase noise performance of the PLL, providing insights into the dominant noise sources and their impact on system behavior.
keywords: 
    - Integer-N PLL
    - Phase Noise Analysis
    - Noise Modeling
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

@@

# Introduction

A `{python} format(f0)`Hz–`{python} format(f)`Hz synthesizer with a step size of `{python} format(step)`Hz is required. Given that the PLL is an Integer-N type, with a reference frequency $f_\text{ref}=$ `{python} format(fref)`Hz, a charge pump delivering `{python} format(I_CP)`A of output current, and a VCO operating from a `{python} format(V_sup)`V supply are available, the objective is to design an Integer-N synthesizer with a loop bandwidth of `{python} format(f_3dB)`Hz using these components. This example is adapted from @rogers2006IntegratedCircuitDesign.

# Design Methodology

## Voltage-Controlled Oscillator (VCO)

Considering a supply voltage of `{python} format(V_sup)`V and a tuning range of `{python} format(tunning_range)`Hz, the VCO gain constant is calculated as $K_\text{VCO}=$ `{python} format(Kvco)`Hz/V. The nominal oscillation frequency is arbitrarily selected as $f_\text{VCO}=$ `{python} format(fvco)`Hz. With this frequency, the corresponding division ratio is $N=$ `{python} N`.

@@

As the output current of the charge pump is known, the charge pump gain can be determined as $K_\theta=$ `{python} format(Ktheta)`A/rad.

@@

The design begins with determining the natural frequency:

@@

By selecting a damping factor $\zeta=$ `{python} zeta`, the natural frequency becomes $f_n=$ `{python} format(f_n)`Hz.

The first capacitor, $C_1$, is calculated as:

@@

Thus, $C_1=$ `{python} format(C_1)`F.

The single resistor in the second-order loop filter is then determined by:

@@

This results in $R_1=$ `{python} format(R_1)`$\Omega$. The second capacitor is set to one-tenth of $C_1$, yielding $C_2=$ `{python} format(C_2)`F.

# Phase Noise Analysis

This section presents the analysis of phase noise in the previously designed Integer-N PLL.

@@

The VCO phase noise is assumed to follow the 20 dB/decade slope, with $PN_\text{VCO}=$ `{python} pn_vco` dBc/Hz at an offset of $\Delta f=$ `{python} format(offset_vco)`Hz. The noise floor is given by $PN_{\text{VCO}_0}=$ `{python} noise_floor` dBc/Hz. Additionally, the charge pump produces a noise current of $i_{n_\text{CP}}=$ `{python} format(I_noise_CP)`A/$\sqrt{\text{Hz}}$. Noise contributions from the PFD, divider, and reference are neglected.

Under the assumption of a 20 dB/decade falloff, the VCO phase noise can be modeled as:

@@

To apply this model, the parameters $C$ and $D$ must be determined.

The coefficient $C$ is given by:

@@

Thus, $C=$ `{python} round(C_param, 1)` $\text{rad}^4/\text{Hz}^2$. The constant term $D$ is computed as:

@@

Resulting in $D=$ `{python} format_e(D_param)` $\text{rad}^2/\text{Hz}$.

@@

The noise current from the charge pump is input-referred by:

@@

Substituting values gives $\varphi_\text{CP}=$ `{python} format(phi_CP)`$\text{rad}/\sqrt{\text{Hz}}$.

## Loop Filter Noise Contribution

To analyze the noise contributed by the second-order loop filter, consider that the noise voltage produces a current through the series combination of $C_1$, $C_2$, and $R$:

@@

Similarly, the voltage analysis yields:

@@

Where the resistor provides the current noise $i_n=\sqrt{4kT/R}$. Referring the voltage noise to the input results in:

@@

## Transfer Function of Input-Referred Noise

The transfer function from input-referred noise to the output is given by:

$$
\begin{split}
\frac{\varphi_\text{no}(s)}{\varphi_\text{nI}(s)}&=\frac{\frac{F(s)K_\text{VCO}K_\theta}{Ns}}{1+\frac{F(s)K_\text{VCO}K_\theta}{Ns}}\cdot N\cdot\frac{s}{s}\implies\\
&=\frac{F(s)K_\text{VCO}K_\theta}{s+\frac{F(s)K_\text{VCO}K_\theta}{N}}
\end{split}
$$

As the output phase noise PSD $\varphi_\text{no}$ is observed before the divider, it is valid to multiply the result by $N$ (or divide by $N^{-1}$).

For a PLL employing a second-order loop filter, this becomes:

@@

Substituting the component values results in the following numerical transfer function:

@@

display(Markdown(f"$$\\varphi_\\text{{no}}(s) = {latex_str}\\cdot {format_raw(phi_CP, 0)}$$ {{#eq-filter-transfer-function}}"))
```

Finally, the phase noise in dBc/Hz is computed as:

@@

# Conclusion

This report presented the design of an Integer-N PLL synthesizer and its corresponding phase noise analysis. Key parameters such as the VCO and charge pump constants, loop filter components, and noise contributions were derived and evaluated. The system was analyzed under the assumption of a 20 dB/decade phase noise roll-off for the VCO and considered primary noise sources including the charge pump and loop filter.
